---
title: "Practical session on fitting distributions"
author: "F. Bertrand"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    toc: true
#    css: style.css
    number_sections: true
    highlight: github
    fig_height: 8
    fig_width: 12
    includes:
      before_body: header.html
      after_body: footer.html
  html_document:
    toc: true
    number_sections: true
    fig_height: 8
    fig_width: 12
    includes:
      after_body: footer.html
  pdf_document:
    toc: true
    number_sections: true
    fig_height: 4
    fig_width: 6
  word_document:
    toc: true
    fig_height: 8
    fig_width: 12
---


# Introduction

## Package installation

Let's start by installing the packages we will need in the following.
```{r}
if(!require(devtools)){install.packages("devtools");library(devtools)}
if(!require(ggmap)){install.packages("ggmap");library(ggmap)}
```

```{r, include=FALSE, eval=FALSE}
if(!require(ggmap)){
  devtools::install_github("dkahle/ggmap", ref = "tidyup")
  library(ggmap)
  } else {
    if(!(packageDescription("ggmap")$RemoteType=="github")){
      devtools::install_github("dkahle/ggmap", ref = "tidyup")
      library(ggmap)}
    }
```

## Recovery of hydrolic data 

The data to be studied were presented in [Dione (1992)](https://tinyurl.com/y564t3jm). 
```{r}
station_data <- read.csv("https://tinyurl.com/y2kosb2y")
longlat <- read.csv("https://tinyurl.com/yy9f9bsf")
```

The `KIDIRA`, `GOURBA`(ssi) and `FADOU`(gou) columns correspond to annual flows between 1903 and 1989. The columns `GOUc` and `KILIc` correspond to complementary hydrolic data.

The `dms2dd` function of the `biogeo` package allows to convert coordinates known as longitude and latitude expressed in degrees, minutes and seconds into decimal degrees.

```{r}
if(!require(biogeo)){install.packages("biogeo");library(biogeo)}
latdd<-with(longlat,dms2dd(LATDEG,LATMIN,0,LATNS))
latdd
min(latdd)
max(latdd)
longdd<-with(longlat,dms2dd(LONGDEG,LONGMIN,0,LONGOE))
min(longdd)
max(longdd)
typesite <- (longlat$NATURE.DU.SITE=="STATIONS PLUVIOMETRIQUES")+1
df_dd_stations <- data.frame(cbind(latdd,longdd,typesite))
```


## Graphic visualization of the position of the measuring stations.

Recovery of geographical data from the Internet.

`bbox` = *bounding box*.

```{r}
min(longdd)
max(longdd)
min(latdd)
max(latdd)
faleme <- get_stamenmap(bbox = c(left = min(longdd)*1.05, bottom = min(latdd)/1.05, right = max(longdd)/1.05, top = max(latdd)*1.05), 
zoom = 8)
```

Several types of land are available.
```{r}
maptypes = c("terrain", "terrain-background", "terrain-labels", 
             "terrain-lines", "toner", "toner-2010", "toner-2011", 
             "toner-background", "toner-hybrid", "toner-labels", 
             "toner-lines", "toner-lite", "watercolor")
```

Retrieve another example of a map terrain type.
```{r}
faleme_toner2011 <- get_stamenmap(bbox = c(left = min(longdd)*1.05, bottom = min(latdd)/1.05, right = max(longdd)/1.05, top = max(latdd)*1.05), zoom = 8, maptype="toner-2011")
```

Adding the stations to the two types of maps of the region.
```{r}
ggmap(faleme) + geom_point(
  aes(x = longdd, y = latdd, color=typesite),
  data = df_dd_stations, colour = c("red","green")[typesite], size = 3
)
```


```{r}
ggmap(faleme_toner2011) + geom_point(
  aes(x = longdd, y = latdd, color=typesite),
  data = df_dd_stations, colour = c("red","green")[typesite], size = 3
)
```

# Adjustments to hydrological data

```{r}
if(!require(fitdistrplus)){install.packages("fitdistrplus");library(fitdistrplus)}
```

## KIDIRA

Représentations des données.
```{r}
plotdist(station_data$KIDIRA) 
```

```{r}
descdist(station_data$KIDIRA, boot = 1000)
```

### Some examples of fitting distributions to data.

#### Log-normal

```{r}
KIDIRA.ln <- fitdist(station_data$KIDIRA, "lnorm")
summary(KIDIRA.ln)
plot(KIDIRA.ln)
```

#### Gumbel

```{r}
if(!require(actuar)){install.packages("actuar");library(actuar)}
KIDIRA.gumbel <- fitdist(station_data$KIDIRA, "gumbel", start=list(alpha=10, scale=10))
summary(KIDIRA.gumbel)
plot(KIDIRA.gumbel)
```

#### Log-logistic

```{r}
KIDIRA.ll <- fitdist(station_data$KIDIRA, "llogis", start = list(shape = 1, scale = 10))
summary(KIDIRA.ll)
plot(KIDIRA.ll)
```

#### Pareto

```{r}
KIDIRA.P <- fitdist(station_data$KIDIRA, "pareto", start = list(shape = 1, scale = 10))
summary(KIDIRA.P)
plot(KIDIRA.P)
```

#### Burr

```{r}
KIDIRA.B <- fitdist(station_data$KIDIRA, "burr", start = list(shape1 = 1, shape2 = 1))
summary(KIDIRA.B)
plot(KIDIRA.B)
```

### Graphical comparison of the fits

By their densities.
```{r}
denscomp(lapply(paste("KIDIRA",c("ln","gumbel","ll","P","B"),sep="."),get), legendtext = c("lognormal","gumbel", "loglogistic", "Pareto", "Burr"))
```

By quantile-quantile diagrams.
```{r}
qqcomp(lapply(paste("KIDIRA",c("ln","gumbel","ll","P","B"),sep="."),get), legendtext = c("lognormal","gumbel", "loglogistic", "Pareto", "Burr"))
```

By their distribution functions.
```{r}
cdfcomp(lapply(paste("KIDIRA",c("ln","gumbel","ll","P","B"),sep="."),get), legendtext = c("lognormal","gumbel", "loglogistic", "Pareto", "Burr"))
```

By probability-probability diagrams.
```{r}
ppcomp(lapply(paste("KIDIRA",c("ln","gumbel","ll","P","B"),sep="."),get), legendtext = c("lognormal","gumbel", "loglogistic", "Pareto", "Burr"))
```

### Numerical comparison of fits

```{r}
gofstat(lapply(paste("KIDIRA",c("ln","gumbel","ll","P","B"),sep="."),get), fitnames = c("lognormal","gumbel", "loglogistic", "Pareto", "Burr"))
```


## Other distributions

The initial study suggests testing the fit to Galton's (= lognormal), Goodrich's, Pearson's 3, Gumbel, ...

## Other techniques

Apply alternative distribution selection approaches to these data.


# Applications to other variables.

```{r}
plotdist(station_data$GOURBA) 
```

```{r}
plotdist(station_data$FADOU) 
```

```{r}
plotdist(station_data$GOUc) 
```

```{r}
plotdist(station_data$KILIc) 
```

```{r}
descdist(station_data$GOURBA, boot = 1000)
```

```{r}
descdist(station_data$FADOU, boot = 1000)
```

Valeurs annuelles cumulées.
```{r}
descdist(station_data$GOUc, boot = 1000)
```

```{r}
descdist(station_data$KILIc, boot = 1000)
```

# Mystery data

Which distribution(s) is/are compatible with this data?

```{r}
donnees <- read.csv("https://tinyurl.com/yxn9udgt")
str(donnees)
```


